---
title: "Fatal Police Victims of 2016"
author: "Dylan Arguelles"
output: html_document
---

```{r, echo=FALSE}
#library(ggplot2)
#library(dplyr)
#library(maps)
```
## 1,093 People were killed by Police in 2016. 92.5% by gunshot. Here is an analysis of the victims.


```{r,echo=FALSE, message=FALSE}
library(readr)
the_counted_2016 <- read_csv("~/Desktop/popo/policeKillings/data/the-counted-2016.csv")

```


## What State has the Most Deadly Police

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(maps)
library(mapproj)
# TODO normalize by population and check frequency

#frequency per state
killstates <- data.frame(count(the_counted_2016,state))
killstates$state <-state.name[match(killstates$state,state.abb)]
killstates$region <- tolower(killstates$state)
states <- map_data("state")

#combine killings & map data frame
map.df <- merge(states,killstates, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]

#plot heatmap of US for killings per state
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=n))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+
  coord_map()+
  labs(title = "Killings by State", fill = "Count")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```




## Are Races Targeted?

# 24.3% of victims are black compared to 12.6% of the total US population
```{r, echo=FALSE, results='hide',fig.keep='all'}
# TODO compare with total population percentages

#plot bar graph 
ggplot(the_counted_2016, aes(x=raceethnicity, color=raceethnicity))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  labs(title = "Killings by Race/Ethnicity",
       x = "Race/Ethnicity",
       y = "Number of Killings")

```


## How Old are the Victims

# Most victims are young to middle aged adults.
# 19 minors were killed.
```{r, echo=FALSE, warning=FALSE}
#format data
the_counted_2016$ageRange <- cut(as.numeric(the_counted_2016$age), breaks=c(0,25,40,60,100), labels=c("<=25","26-40","41-60",">60"))


#plot bar graph
ggplot(the_counted_2016, aes(x=ageRange)) + geom_bar(fill="steelblue")


```

## Gender Roles

# Victims are overwhelmingly male
```{r, echo=FALSE, message=FALSE}
#format data
g <- data.frame(count(the_counted_2016,gender))
g$p <- scales::percent(g$n/sum(g$n))

#plot pie chart
ggplot(g, aes(x="", y= n, fill=gender))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0)+
  theme_void()+
  theme(axis.text.x=element_blank())+
  geom_text(aes(x=1, y = cumsum(n) - n/2, label=p))+
  labs(title = "Killings by Gender")



```

## What Time of the Year is the Most Deadly

# There doesn't appear to be a trend with the seasons
```{r, echo=FALSE, message=FALSE}
ggplot(the_counted_2016, aes(x=factor(the_counted_2016$month, levels = month.name), 
  color=month))+
  geom_bar()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(title = "Killings by Month",
       x = "Month",
       y = "Number of Killings")


```
